### 2014-10-24 paper/experiments.tex
adding some changes
adding changes

We conducted several performance tests to measure the throughput and scalability of our system. All these tests were performed in an LAN system called Lattice which has a network bandwidth of 1Gbps. All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory. Sample code used for all tests can be found here.

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. For our solution and Yahoo S4! EventProducer reads a file contains over 7500000 ecg records and push events with multiple threads. Twitter Storm does not allow user threads to push data. So we use the Spout thread to send messages into the system. EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.  
We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case. Throughput was measured at the EventProducer calculating total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

@@

As shown in the Figure \ref{throuput} our solution out performs Twitter storm and Yahoo S4!. However by looking the Figure \ref{throuput} one might think even our system does not scalable although it perform high. In order to examine the reason behind this we need to look into the network bandwidth and CPU load average values.

@@

As shown in the figure \ref{efficiency} Twitter storm uses a minimum over head while Yahoo S4! performs badly compared to other two. This overhead is due to its usage of  java serializations to serialize data. Then we analyzed to which our solution spends extra 32 bytes. Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server. Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and other 4 bytes due to tcp over head. This is the trade off we had to pay for improve the parallelism compared to direct one tcp communication per process.

@@

This application predicts load averages using the previous values and a given machine learning algorithm. We implement this logic using nodes as shown in the Figure \ref{multigraph}. First producer reads the data file and send events to the avgCal node which calculate the last minite average and send the same event to both plugPredict and housePredict processors. Both plugPredict and housePredict processors predict the next values and send events to receivers. The original problem only requires to send those prediction events in 30s intervals. But we send a prediction event fore each messages to observe how the system works with high load.

>>>>

We conducted several performance tests to measure the throughput and scalability of our system. All these tests were performed in an LAN system called Lattice which has a network bandwidth of 1Gbps. All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 1 Gbp of memory. Sample code used for all tests can be found here\cite{solutionCode}.

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. For our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file contains over 7500000 ecg records and push events with multiple threads. Twitter Storm\cite{twitterStorm} does not allow user threads to push data. So we used the Spout thread to send messages into the system. EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.  
We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case. Throughput was measured at the EventProducer calculating total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

@@

As shown in the Figure \ref{throuput} our solution out performs Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}. However by looking the Figure \ref{throuput} one might think even our system does not scalable although it perform high. In order to examine the reason behind this we need to look into the network bandwidth and CPU load average values.

@@

As shown in the figure \ref{efficiency} Twitter storm\cite{twitterStorm} uses a minimum over head while Yahoo S4\cite{neumeyer2010s4} performs badly compared to other two. This overhead is due to its usage of  java serializations to serialize data. Then we analyzed to which our solution spends extra 32 bytes. Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process and to parse the message at the server. Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes to each message at application level and other 4 bytes due to tcp over head. This is the trade off we had to pay for improve the parallelism compared to direct one tcp communication per process.

@@

This application predicts load averages using the previous values and a given machine learning algorithm. We implement this logic using nodes as shown in the Figure \ref{multigraph}. First producer reads the data file and send events to the avgCal node which calculate the last minite average and send the same event to both plugPredict and housePredict processors. Both plugPredict and housePredict processors predict the next values and send events to receivers. The original problem only requires to send those prediction events in 30s intervals. But we send a prediction event for each messages to observe how the system works with high load.

### 2014-10-24 paper/relatedwork.tex
adding changes

Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the populariy of Apache Hadoop\cite{hadoop}. In a map reduce envirionment a problem is partition into smaller parts identified by a key and process parallely. Although these systems support user defined functions, data processing flow of these systems are fixed. Ciel\cite{zaharia2012resilient} address this problem by dynamically generating dataflow graph. Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.

@@

Spark\cite{boyd2011distributed} introduces its RRDs\cite{zaharia2012resilient} to achive better fault tolarance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically. Spark Streaming\cite{murray2013naiad} extends this concept to distributed stream processing with the concept of discretized events. Spark streaming\cite{murray2013naiad} processes data as batches while storing them as RDDs \cite{agarwal2011reliable} to achive higher throughput. This batch processing may introduce latances which are not acceptable for some applications. Further it is not clear how to implement complex event processing algorithumns such as we use our bench marks with Spark Streaming. MillWheel\cite{zaharia2013discretized} is the stream processing system used at Google to process web queries. MillWheel provides relaiable message processing as well as state persistance for its graph based computations. Unlike in our system MillWheel treat messages as binary messages leaving message parsing and serialization to application layer.

>>>>

Map reduce\cite{dean2008mapreduce} is a widely adapted technology to process batch data with the populariy of Apache Hadoop\cite{hadoop}. In a map reduce envirionment a problem is partition into smaller parts identified by a key and process parallely. Although these systems support user defined functions, data processing flow of these systems are fixed. Ciel\cite{murray2011ciel} address this problem by dynamically generating dataflow graph. Since all these systems communicate through file system they inherently not suitable for real time processing\cite{lam2012muppet}.

@@

Spark\cite{zaharia2010spark} introduces its RRDs\cite{zaharia2012resilient} to achive better fault tolarance by storing the operations and replaying them instead of replicating data itself or use checkpointing to store data periodically. Spark Streaming\cite{zaharia2013discretized} extends this concept to distributed stream processing with the concept of discretized events. Spark streaming\cite{zaharia2013discretized} processes data as batches while storing them as RDDs \cite{zaharia2012resilient} to achive higher throughput. This batch processing may introduce latances which are not acceptable for some applications. Further it is not clear how to implement complex event processing algorithumns such as we use our bench marks with Spark Streaming. MillWheel\cite{akidau2013millwheel} is the stream processing system used at Google to process web queries. MillWheel\cite{akidau2013millwheel} provides relaiable message processing as well as state persistance for its graph based computations. Unlike in our system MillWheel treat messages as binary 
messages leaving message parsing and serialization to application layer.

### 2014-10-24 paper/system_architecture.tex
adding changes

\section{System Design}
In this section we elaborate the underling system design which is capable of processing 2.4 million messages per second. We emphazise three features of this design. Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages. All processes share the same set of connections which we called as the bridge. The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we show latter our experiments shows this approach perform several times better than the other systems dispite above overhead. Secondly we have implemented java.io.OutputStream and java.io.InputStream interfaces on top of java.nio API. Using these input and output streams we provide java.io.DataOutput and java.io.DataInput 
objects to directly serialize message data to binary format to without sending any metadata along the message. As we explained in earlier section our Event interface provides these methods. Finally except messages itself all the other objects used to send the messages and parse the messages are not created at message processing time. When a process sends the first message all these objects are created and it is a one time cost. Rest of this section further explain how we have implemented above techniques.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involves in sending a message from one process to another. Lets assume there is a message object at a client side process and that process wants to send this message to another process. First client process sends this message to its ElementContainer. ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream object decides which node to send this message and pass the message along the taget node details to ConnectionManager which holds ClientConnection object for each node. ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection. ClientConnection picks an available DataOutput object and invokes the serialize method of the message to send that to server side.

At the server side there are a set of ServerTasks which read receiving messages at DataInput streams available in the ServerConnection (we register these connections at the connections creating time). Once a server task receives a binary message, it creates the event object using the sending process id to identify the event type. Then it pass this message to the WorkerContainer which holds all processes. Finally WorkerContainer dispatch this message to correct process uisng receiving process id. Message ordering happens at the process level if required. This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream without creating objects dynamically. 
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented java.io.OutputStream interface using our DataWriter class. DataWriter contains a java.nio.ByteBuffer object. When a higher layer method writes data to DataOuput object, DataWriter put those bytes to its byte buffer. When a selection event occurs, Datawriter writes its byte buffer to under line SocketChannel. This way we use one byte buffer as a intermediatory create an output stream with java.nio API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented java.io.InputStream interface using our DataReader class. Similar to clinet side, DataReader object has a java.nio.ByteBuffer object. When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer. When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with java.nio API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}

>>>>

\section{System Design}
In this section we elaborate the underling system design which is capable of processing 2.4 million messages per second. We emphazise three features of this design. Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages. All processes share the same set of connections which we called as the bridge. The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we show latter our experiments shows this approach perform several times better than the other systems dispite above overhead. Secondly we have implemented OutputStream and InputStream interfaces on top of java.nio API. Using these input and output streams we provide DataOutput and DataInput APIs to directly serialize message data to binary format without sending any metadata along the message. Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time. When a process sends the first message all these objects are created and it is a one time cost. Rest of this section further explain design techniques we used to realise these concepts.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involves in sending a message from one process to another. Lets assume there is a message at a client side process and that process wants to send this message to another process. First client process sends this message to its ElementContainer. ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream decides which node to send this message and passes the message along the taget node details to ConnectionManager which holds ClientConnections for each node. ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection. ClientConnection picks an available DataOutput from its pool of DataOutputs and invokes the serialize method of the message to send that to server side.

At the server side there is a set of ServerTasks which read receiving messages from pool of DataInput streams available in the ServerConnection (we register these connections at the connections creating time). Once a server task receives a binary message, it creates the event using the sending process id to identify the event type. Then it pass this message to the WorkerContainer which holds all processes. Finally WorkerContainer dispatch this message to correct process uisng receiving process id. Message ordering happens at the process level if required. This elastrates how we have shared a pool of tcp connections among process and next we show how we have implemented java.io.OutputStream and java.io.InputStream with java.nio API without  creating objects dynamically.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented java.io.OutputStream interface using our DataWriter. DataWriter contains a ByteBuffer. When a higher layer method writes data to DataOuput, DataWriter put those bytes to its byte buffer. When a selection event occurs, Datawriter writes its byte buffer to underline SocketChannel. This way we use one byte buffer as an intermediatory to create an output stream with java.nio API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented java.io.InputStream interface using our DataReader. Similar to clinet side, DataReader has a java.nio.ByteBuffer. When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer. When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with java.nio API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}

### 2014-10-24 paper/system_overview.tex
adding changes

\section{System Overview}
Our system mainly consists of a set of processes which receive events, process them and forward to next processes. We use the term adapters for special processes which receive events from out side systems. These processes and adapters forms a graph which represents the way a given event stream being processed within the system. Rest of this section further describes this concept at a user level. Our current implementation does not support fault tolarance and have adapted topoloy concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}. In the next section we provide a detail explanation about how we have make this interprocess communication efficient.
\subsection{Process Graph}
A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which does the processing logic and their interaction patterns. Adapters receive events from out side systems and forward them to other processes. Processes process events according to a given logic and emits the generated events either to other processes or out side systems.

@@

\subsection{Specifying the Runtime Graph}

A runtime graph mainly consists of a set of processes, adapters and events. Processes, adapters and events can be created by implementing respective interfaces as shown below.

@@

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,    
  tabsize=2
}

\begin{lstlisting}
public interface Event {
    public String getKey();
    public void serialize(DataOutput dataOutput);
    public void parse(DataInput dataInput);
}

public interface Element {
    public void initialise(Container container,
                     Map<String,String> parameters);
}

public interface Adaptor extends Element {
    public void start();
}

public interface Processor extends Element {
    public void onEvent(Event event);
}
\end{lstlisting}

 The Event interface has a method to get the key for a given event. This key is used in sending messages to next processes as described earlier. Parse and serialize methods of event interface are used to direcly serialize the object state to underline streams. Process interface has a method called onEvent which underline framework invokes when it receives an event to that process. The start method of the adapter interface is invoked by the framework and that can be used to pull events from out side systems onces the system starts. At the initilization time both processes and adapters receive the container object which can be used to send events to other processes. Users can specify the runtime graph to the system using a json file as shown below.

@@

A deployment of the system consists of a manager node and a set of worker nodes \ref{deployment}. Worker nodes are grouped in to clusters so that at the runtime graph users can specify to which clusters each processes needs to get deployed. Once the system starts worker nodes communicate with each other passing messages.

>>>>

\section{System Overview}
Our system mainly consists of a set of processes which receive events, process them and forward to next processes. We use the term adapters for special processes which receive events from out side systems. These processes and adapters forms a graph which represents the way a given event stream being processed within the system. Our current implementation does not support fault tolarance and have adapted topoloy concept from twitter storm\cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}. Rest of this section further describes this concept at a user level. In the next section we provide a detail explanation about how we have make this interprocess communication efficient. 
\subsection{Process Graph}
A Process graph (Figure \ref{processgraph}) mainly consists of the adapters, processes which does the processing logic and their interaction patterns. Adapters receive events from out side systems and forward them to other processes. Processes process events according to a given logic and emits the generated events either to other processes or to out side systems.

@@

\subsection{API}
In this section, we explain our API which can be used to define custom events, processes and adapters. We provide clear interfaces to abstract out all the communication complexities from application developers.

@@

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  frame=single,
  breaklines=true,
  breakatwhitespace=true,    
  tabsize=2
}

\begin{lstlisting}
public interface Event {
    String getKey();
    void serialize(DataOutput dataOutput);
    void parse(DataInput dataInput);
}

public interface Element {
    void initialise(Container container,
            Map<String,String> parameters);
}

public interface Adaptor extends Element {
    void start();
}

public interface Processor extends Element {
    void onEvent(Event event);
}
\end{lstlisting}

 The Event interface has a method to get the key for a given event. This key is used in sending messages to next processes as described earlier. Parse and serialize methods of event interface are used to direcly serialize the event attributes to underline streams. When implementing these methods deveopers can define attibute serializing order in serialize method and use same order at the parse method to avoid meta data passing with the binary format. Process interface has a method called onEvent which underline framework invokes when it receives an event to that process. The start method of the adapter interface is invoked by the framework and that can be used to pull events from out side systems onces the system starts. At the initilization time both processes and adapters receive the container which can be used to send events to other processes.

As shown in Figure \ref{deployment}, a deployment of the system consists of a manager node and a set of worker nodes. Worker nodes are grouped into clusters so that the application users can specify to which clusters each processes needs to get deployed. Manager node process applications, deploy them to worker nodes and initiate the processing by starting the adapters.

\subsection{Application Developement}

An application specifies the runtime graph to process a pirticular event stream. Therefore an appication mainly consists of a set of processes, adapters and events and their interactions. Processes, adapters and events can be developed by implementing the respective interfaces as shown above. Then the interactions can be specified using a json file.

@@

As shown above, the \textit{instances} parameter can be used to specify the required number of instances to be deployed. The \textit{receivers} parameter is used specify from which process it expects to receive messages and how the sending process should distribute the events among instances. The \textit{key} type indicates that events with same key must be received for same instance. Finally users can deploy the application to manager which process them and deploys to workers.
